---
title: "bean Atlas maps"
author: "A Farrow"
date: "04/10/2021"
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage[T1]{fontenc}
mainfont: SourceSansPro
output:
  pdf_document:
    dev: png
fontsize: 12pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  dpi = 300,
  echo = FALSE,
  results = 'hide' ,
  fig.path = "figures/",
  dev = 'png',
  message = FALSE,
  warning = FALSE
)

options(width = 250, dplyr.width = 120)

```

```{r}

# Set so that long lines in R will be wrapped:
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60), tidy=TRUE)

```

```{r packages}

library(terra)
library(raster)
library(sf)
library(tidyverse)
library(ggplot2)
library(basemaps)
library(ggmap)
library(foreign)
library(RColorBrewer)
library(naniar)
library(gridExtra)

```

# Introduction

Here I load the data from the first and second bean atlases of africa.

# Load data

## Basemap

There are two types of basemaps:

(1) GADM country boundaries
(2) terrain mapping with oceans


``` {r, results = 'asis', echo = T, cache=TRUE, fig.height = 4, fig.width = 7, fig.cap = "GADM country boundaries", dev = 'png'}


misc <- list()
misc$countries <-
  c(
    "AGO",
    "BDI",
    "BEN",
    "BFA" ,
    "BWA",
    "CAF" ,
    "CIV" ,
    "CMR" ,
    "COD" ,
    "COG" ,
    "COM" ,
    "CPV" ,
    "DJI" ,
    "DZA" ,
    "EGY" ,
    "ERI" ,
    "ESH",
    "ETH",
    "GAB" ,
    "GHA",
    "GIN" ,
    "GMB" ,
    "GNB" ,
    "GNQ" ,
    "KEN" ,
    "LBR",
    "LBY" ,
    "LSO" ,
    "MAR" ,
    "MDG" ,
    "MLI" ,
    "MOZ" ,
    "MRT" ,
    "MUS" ,
    "MWI" ,
    "NAM" ,
    "NER" ,
    "NGA" ,
    "REU" ,
    "RWA" ,
    "SDN" ,
    "SEN" ,
    "SLE" ,
    "SOM" ,
    "SSD",
    "STP" ,
    "SWZ" ,
    "TCD" ,
    "TGO" ,
    "TUN" ,
    "TZA" ,
    "UGA" ,
    "ZAF" ,
    "ZMB" ,
    "ZWE"
  )
v_adm_afr <- st_as_sf(do.call("bind", lapply(misc$countries, 
                                    function(x) getData('GADM', country=x, level=0))))


myAfricaMap <- ggplot() +
  geom_sf(
    data = v_adm_afr,
    col = "grey",
    fill = "white",
    alpha = 1,
    fill = NA,
    inherit.aes = FALSE
  )
myAfricaMap

```


``` {r, results = 'asis', echo = T, cache=TRUE, fig.height = 4, fig.width = 7, fig.cap = "Terrain basemap", dev = 'png'}

# basemap with ggmap

myMap <- get_stamenmap(bbox = c(left = -20,
                                    bottom = -36,
                                    right = 52,
                                    top = 23),
          maptype = "terrain-background", 
          crop = T,
          zoom = 6)


# plot map
ggbackground <- ggmap(myMap) 

ggbackground

```

## V1 Bean Atlas polygons

```{r bean_atlas_1, cache=TRUE, fig.height = 4, fig.width = 7, fig.cap = "Version 1 MBPAs - terrain", dev = 'png'}

v_bean_atlas_poly_1 <- st_read("D:/DatosProyecto/SpatialData/Africa/africa/RootRot/beanatlas/Ciatdb/BEANDB/Themes/Beanplys.shp")

st_crs(v_bean_atlas_poly_1) = 4326
v_bean_atlas_poly_1

ggmap(myMap) +
  
  geom_sf(
    data = v_bean_atlas_poly_1,
    aes(fill = BEANPOLY_I),
    col = "red",
    alpha = 0.25,
    inherit.aes = FALSE
  ) 

```


### Attributes 

There are three attribute files in dbf format:

```{r bean_atlas_1_join, cache=TRUE}

# main_crop - intercrop dbf details

dB_bean_atlas_main_crop <- read.dbf("D:/DatosProyecto/SpatialData/Africa/africa/RootRot/beanatlas/Ciatdb/BEANDB/Tables/Main_crop.dbf", as.is = TRUE)

str(dB_bean_atlas_main_crop)

v_bean_atlas_poly_1_main_crop <- inner_join(v_bean_atlas_poly_1, dB_bean_atlas_main_crop, by = c("BEANPOLY_I" = "PLY2_ID"))
str(v_bean_atlas_poly_1_main_crop)


# main_var - variety diveristy details

dB_bean_atlas_main_var <- read.dbf("D:/DatosProyecto/SpatialData/Africa/africa/RootRot/beanatlas/Ciatdb/BEANDB/Tables/Main_var.dbf", as.is = TRUE)

str(dB_bean_atlas_main_var)

v_bean_atlas_poly_1_main_var <- inner_join(v_bean_atlas_poly_1, dB_bean_atlas_main_var, by = c("BEANPOLY_I" = "PLY2_ID"))
str(v_bean_atlas_poly_1_main_var)


# afptdata

dB_bean_atlas_afptdata <- read.dbf("D:/DatosProyecto/SpatialData/Africa/africa/RootRot/beanatlas/Ciatdb/BEANDB/Tables/afptdata.dbf", as.is = TRUE)

str(dB_bean_atlas_afptdata)


# main_mbp - general attributes

dB_bean_atlas_main_mbp <- read.dbf("D:/DatosProyecto/SpatialData/Africa/africa/RootRot/beanatlas/Ciatdb/BEANDB/Tables/Main_mbp.dbf", as.is = TRUE)

str(dB_bean_atlas_main_mbp)

v_bean_atlas_poly_1_main_mbp <- inner_join(v_bean_atlas_poly_1, dB_bean_atlas_main_mbp, by = c("BEANPOLY_I" = "PLY2_ID"))
str(v_bean_atlas_poly_1_main_mbp)

```

Map the joins


```{r bean_atlas_1_join_map_1, cache=TRUE, fig.height = 4, fig.width = 7, fig.cap = "Version 1 MBPAs - GADM", dev = 'png'}

myAfricaMap +
  geom_sf(
    data = v_bean_atlas_poly_1_main_mbp,
    aes(fill = MBPA_ID),
    col = "red",
    alpha = 0.25,
    inherit.aes = FALSE
  ) +
  geom_sf_text(
    data = v_bean_atlas_poly_1_main_mbp,
    stat = "sf_coordinates",
    position = "identity",
    aes(geometry = geometry, label = paste(REGION)),
    size = 2,
    inherit.aes = FALSE
  )

```

```{r bean_atlas_1_join_map_2, cache=TRUE}

maize_1 <- myAfricaMap +  
  geom_sf(
    data = v_bean_atlas_poly_1_main_crop,
    aes(fill = MZ),
    col = NA,
    alpha = 1,
    inherit.aes = FALSE
  )  + 
    scale_fill_distiller(direction = -1, trans = "reverse", type = "seq", palette = "Greens")


sole_1 <- myAfricaMap +  
  geom_sf(
    data = v_bean_atlas_poly_1_main_crop,
    aes(fill = SC),
    col = NA,
    alpha = 1,
    inherit.aes = FALSE
  )  + 
    scale_fill_distiller(direction = -1, trans = "reverse", type = "seq", palette = "Blues")


habit_1 <- myAfricaMap +
  geom_sf(
    data = v_bean_atlas_poly_1_main_mbp,
    aes(fill = HABIVPRE),
    col = NA,
    alpha = 1,
    inherit.aes = FALSE
  ) +
    scale_fill_manual(values = c( "light green", "light yellow", "sky blue","dark green")) +
  labs(fill = "----------\nC Bean\n----------")
 
```

## V2 Bean Atlas polygons

```{r bean_atlas_2, cache=TRUE, fig.height = 4, fig.width = 7, fig.cap = "Version 2 MBPAs", dev = 'png'}

v_bean_atlas_poly_2 <- st_read("D:/DatosProyecto/SpatialData/Africa/africa/RootRot/beanatlas/Ciatdb/BEANDB/Themes/Dataverse/AtlasBean_Data.shp")

st_crs(v_bean_atlas_poly_2) = 4326
str(v_bean_atlas_poly_2)


myAfricaMap +
    geom_sf(
    data = v_bean_atlas_poly_2,
    aes(fill = MBPA_ID),
    col = "red",
    alpha = 0.25,
    inherit.aes = FALSE
  )  +
  geom_sf_text(
    data = v_bean_atlas_poly_2,
    stat = "sf_coordinates",
    position = "identity",
    aes(geometry = geometry, label = paste(MBPA_ID)),
    size = 2,
    inherit.aes = FALSE
  )

```

Plot variables

```{r bean_atlas_2_plot, cache=TRUE}

# maize intercrop

v_bean_atlas_poly_2NA <- v_bean_atlas_poly_2 %>% replace_with_na(replace = list(CS_Maize = c(-999, -9999)))

str(v_bean_atlas_poly_2NA)
maize_2 <- myAfricaMap +  
  geom_sf(
    data = v_bean_atlas_poly_2NA,
    aes(fill = CS_Maize),
    col = NA,
    alpha = 1,
    inherit.aes = FALSE
  )  + 
    scale_fill_distiller(direction = -1, trans = "reverse", type = "seq", palette = "Greens")


# sole crop

v_bean_atlas_poly_2NA <- v_bean_atlas_poly_2NA %>% replace_with_na(replace = list(CS_Sole = c(-999, -9999)))

sole_2 <- myAfricaMap +  
  geom_sf(
    data = v_bean_atlas_poly_2NA,
    aes(fill = CS_Sole),
    col = NA,
    alpha = 1,
    inherit.aes = FALSE
  )  + 
    scale_fill_distiller(direction = -1, trans = "reverse", type = "seq", palette = "Blues")


# climbing bean

v_bean_atlas_poly_2NA <- v_bean_atlas_poly_2 %>% replace_with_na(replace = list(SD_GH_CL_A = c(-999, -9999)))

habit_2 <- myAfricaMap +  
  geom_sf(
    data = v_bean_atlas_poly_2NA,
    aes(fill = SD_GH_CL_A),
    col = NA,
    alpha = 1,
    inherit.aes = FALSE
  )  + 
    scale_fill_distiller(direction = -1, trans = "reverse", type = "seq", palette = "YlGn", na.value="white") +
  labs(fill = "----------\nC Bean\n----------")


# semi-climbing bean

v_bean_atlas_poly_2NA <- v_bean_atlas_poly_2 %>% replace_with_na(replace = list(SD_GH_SCL_ = c(-999, -9999)))

habit_2s <- myAfricaMap +  
  geom_sf(
    data = v_bean_atlas_poly_2NA,
    aes(fill = SD_GH_SCL_),
    col = NA,
    alpha = 1,
    inherit.aes = FALSE
  )  + 
    scale_fill_distiller(direction = -1, trans = "reverse", type = "seq", palette = "YlGn", na.value="white") +
  labs(fill = "----------\nS Bean\n----------")


```

# Compare

Here I plot the maps from the first and second versions together for a visual analysis of changes.


## Sole crop

```{r sole_compare, cache=TRUE, fig.height = 10, fig.width = 7, fig.cap = "Comparison of sole cropping system", dev = 'png'}

grid.arrange(sole_1, sole_2)

```

## Maize intercrop

```{r maize_compare, cache=TRUE, fig.height = 10, fig.width = 7, fig.cap = "Comparison of bean/maize intercropping system", dev = 'png'}

grid.arrange(maize_1, maize_2)

```

## Climbing bean

```{r CB_compare, cache=TRUE, fig.height = 15, fig.width = 7, fig.cap = "Comparison of climbing bean distribution", dev = 'png'}

grid.arrange(habit_1, habit_2, habit_2s)

```